---
title: "Assignment 2 ML"
output:
  word_document: default
  pdf_document: default
---
getwd()

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
library(class)
library(dplyr)
library(tidyverse)
library(ISLR)
library(dummies)
library(caret)


## R Markdown
```{r}

Myfile <- read.csv("UniversalBank.csv")

```
#transform categorical predictors with more than two categories into dummy variables first
```{r}
library(dummies)
library(dplyr)
library(class)
Myfile$Education <-as.factor(Myfile$Education)
#remove ID and ZIP 
df= dummy.data.frame (select(Myfile, -c(ZIP.Code,ID)))
df$Personal.Loan <-as.factor(df$Personal.Loan)
df$CCAvg <-as.integer(df$CCAvg)
df

```
#remove ID & Zip Code
#partion the data (60% training and validation 40%)
```{r}
library(caret)
library(FNN)
library(class)
library(dplyr)
library(ISLR)

set.seed(1)
train_index = sample (row.names(df),0.6*dim(df)[1])
test.index <- setdiff(row.names(df),train_index)

train_Data = df[train_index,]
Validation_Data= df[test.index,]

#Constraints
Constraints =data.frame(Age =as.integer(40), Experience =as.integer(10), Income = as.integer(84), Family = as.integer(2), CCAvg =as.integer(2), Education1 =as.integer(0), Education2 =as.integer(1), Education3 = as.integer(0), Mortgage = as.integer(0), Securities.Account = as.integer(0), CD.Account = as.integer(0), Online = as.integer(1), CreditCard = as.integer(1))

normal <- preProcess(train_Data[,-c(10)], method=c("center","scale"))   
normal
train_Data <- predict(normal,train_Data)
Validation_Data <- predict(normal,Validation_Data)
Constraints <- predict(normal,Constraints)

K1 <-knn(train=train_Data[,-c(10)],test=Constraints,cl=train_Data[,10], k=1, prob = TRUE)
knn.attributes <- attributes(K1)
knn.attributes[3]
```
#choice of k that balances overfitting and ignoring the predictor

```{r}
bestk <- data.frame(k = seq(1,10,1), accuracy = rep(0,10))

for(i in 1:10) {
  K2 <- knn(train = train_Data[,-10],test = Validation_Data[,-10], cl = train_Data[,10], k=i, prob=TRUE)
  bestk[i, 2] <- confusionMatrix(K2, as.factor(Validation_Data$Personal.Loan))$overall[1] # loop to generate accuracy for k values from 1 to 10
}
bestk

```


#3 Confusion matrix 
```{r}
library(caret)
library(class)
K2 <-knn(train=train_Data[,-10],test=Validation_Data[,-10],cl=train_Data[,10], prob = TRUE)
confusionMatrix(K2, as.factor(Validation_Data$Personal.Loan))


```
`

#Consider the following customer: Age = 40, Experience = 10, Income = 84,Family = 2, CCAvg = 2, Education_1 = 0, Education_2 = 1, Education_3 = 0, Mortgage = 0, Securities Account = 0, CD Account = 0, Online = 1 and Credit Card = 1. Classify the customer using the best k.
```{r}
Customer.df =data.frame(Age =40, Experience = (10), Income = (84), Family = (2), CCAvg = (2), Education1 = (0), Education2 = (1), Education3 = (0), Mortgage = (0), Securities.Account = (0), CD.Account = (0), Online = (1), CreditCard = (1))
K4 <- knn(train = train_Data[,-10],test = Customer.df, cl = train_Data[,10], k=3, prob=TRUE)
K4
```

#Repartion the data (50% training and validation 30% and 20%)
```{r}

library(caret)
set.seed(100)
#partition 50% of training
train_index = sample (row.names(df),0.5*dim(df))
#partition 30% validations
valid.index <- sample(setdiff(rownames(df),train_index),0.3*dim(df))
test.index <- setdiff(row.names(df),train_index)


train_Data = df[train_index,]
Validation_Data= df[valid.index,]
Test_data=df[test.index,]
                
normal <- preProcess(train_Data[,-c(10)], method=c("center","scale"))   
normal
train_Data[,-c(10)] <- predict(normal,train_Data[,-c(10)] )
Validation_Data[,-c(10)] <- predict(normal,Validation_Data[,-c(10)])
Test_data[,-c(10)] <- predict(normal, Test_data[,-c(10)])
Constraints <- predict(normal,Constraints)



K3T <-knn(train=train_Data[,-c(10)],test=train_Data[,-c(10)],cl=train_Data[,10], k=5, prob = TRUE)
K4V <-knn(train=train_Data[,-c(10)],test= Validation_Data[,-c(10)],cl=train_Data[,10], k=5, prob = TRUE)
K5TT <-knn(train=train_Data[,-c(10)],test=Test_data[,-c(10)],cl=train_Data[,10], k=5, prob = TRUE)
confusionMatrix(K3T,train_Data[,10])
confusionMatrix(K4V,Validation_Data[,10])
confusionMatrix(K5TT,Test_data[,10])
```

