---
title: "Assignment 3"
output:
  word_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Packages loaded
library(class)
library(dplyr)
library(tidyverse)
library(ISLR)
library(dummies)
library(caret)
library(dcast)
library(pivottabler)
library(reshape)
library(e1071)
library(naivebayes)
library(klaR)
library(bnclassify)
library(rmarkdown)
library(tinytex)



#Reading the Universal Bank file 
```{r}
Myfile <- read.csv("UniversalBank (1).csv")
Myfile 
head(Myfile)

Myfile$Personal.Loan =as.factor(Myfile$Personal.Loan)
Myfile$CCAvg =as.factor(Myfile$CCAvg)
Myfile$Online =as.factor(Myfile$Online)
```

#Partition the data into training (60%) and validation (40%) sets.
```{r}
set.seed(1)
train_index = sample(row.names(Myfile),0.6*dim(Myfile))
test.index <- setdiff(row.names(Myfile),train_index)

train_Data = Myfile[train_index,]
Validation_Data= Myfile[test.index,]
head(train_Data)
```

#Create a pivot table for the training data with Online as a column variable,CC as a row variable, and Loan as a secondary row variable 
```{r}
library(maditr)
library(reshape)
library(reshape2)
library(ggplot2)
melted.bank = melt(train_Data,id=c("CreditCard" , "Personal.Loan"), variable= "Online")
recast.bank=dcast(melted.bank,CreditCard+Personal.Loan~Online)
recast.bank[,c(1:2,14)]
```


#Consider the task of classifying a customer who owns a bank credit card and is actively using online banking services. Looking at the pivot table, what is the probability that this customer will accept the loan offer? [This is the probability of loan acceptance (Loan = 1) conditional on having a bank credit card (CC = 1) and being an active user of online banking services (Online = 1)].
```{r}
#Probability of customer accepting loan offer= .025 = 2.56%
77/(1924+198+801+77)
```

#Create two separate pivot tables for the training data. Onewill have Loan (rows) as a function of Online (columns) and the other will have Loan (rows) as a function of CC.
```{r}
#pivot table for training data that has loan as a function of Online
melted.bank1 = melt(train_Data,id=c("Personal.Loan"),variable = "Online")
recast.bank1=dcast(melted.bank1,Personal.Loan~Online)
Loan=recast.bank1[,c(1,13)]
Loan

#pivot table for training data that has loan as a function of CC
melted.bank2 = melt(train_Data,id=c("Personal.Loan"),variable = "CreditCard")
recast.bank2=dcast(melted.bank2,Personal.Loan~CreditCard)
LoanCC = recast.bank2[,c(1,14)]
LoanCC

```

#Compute the following quantities [P(A | B) means “the probability ofA given B”]: i.P(CC = 1 | Loan = 1) (the proportion of credit card holders among the loan acceptors)ii.P(Online = 1 | Loan = 1) iii.P(Loan = 1) (the proportion of loan acceptors) iv.P(CC = 1 | Loan = 0) v.P(Online = 1 | Loan = 0)vi.P(Loan = 0)

#i
```{r}
#P(CC = 1 | Loan = 1) 
table(train_Data[,c(14,10)])

#Probability of P(CC = 1 | Loan = 1) = 77/(77+198) = .28 = 28 % is the probability of the Credit Card being 1 and the Personal loan being 1 
77/(77+198)
```
 

#ii
```{r}
#P(Online = 1 | Loan = 1)
table(train_Data[,c(13,10)])

#P(Online = 1 | Loan = 1) = 166/(166+109) = .6036 = 60.36% 
166/(166+109)
```

#iii
```{r}
#P(Loan = 1) (the proportion of loan acceptors)
table(train_Data[,c(10)])

#P(Loan = 1)= 279/ (279+2721) = .091 = 9.1% 
275/ (275+2725)
```

#iv
```{r}
#P(CC = 1 | Loan = 0)
table(train_Data[,c(14,10)])

#P(CC = 1 | Loan = 0) =  801/ (801+1924) = .2939 = 29.39% 
801/ (801+1924)
```

#v
```{r}
#P(Online = 1 | Loan = 0)
table(train_Data[,c(13,10)])
#P(Online = 1 | Loan = 0) = 1588/ (1588+1137) = .5827 = 58.27%
1588/ (1588+1137)
```

#vi
```{r}
#P(Loan = 0)
table(train_Data[,c(10)])
#P(Loan = 0) = 2725 / (2725+275) = .908 = 90.8% 
2725 / (2725+275)
```

#Use the quantities computed above to compute the naive Bayes probability P(Loan = 1 | CC = 1, Online = 1).
```{r}
((77/(77+198)) * (166/(166+109)) * (275/(275+2725))) / (((77/(77+198)) * (166/(166+109)) * (275/(275+2725))) + ((801/(801+1924)) * (1588/(1588+1137)) * 2725/(2725+275)))

#probability is .0905 = 9.05% 
```


#Compare this value with the one obtained from the pivot table in (B). Which is a more accurate estimate?
```{r}
#Comparing the value above, 9.05 % to the value I found in the pivot table which was 2.56%, I think the naive bayes calculation done above is more accurate estimate. 
```

#Which of the entries in this table are needed for computing P(Loan = 1 | CC = 1, Online = 1)? Run naive Bayes on the data. Examine the model output on training data, and find the entry that corresponds to P(Loan = 1 | CC = 1, Online = 1). Compare this to the number you obtained in (E).
```{r}
library(e1071)
library(naivebayes)
library(dplyr)


naive.traindata = train_Data[,c(10,13:14)]
naivebayes= naiveBayes(Personal.Loan~.,data=naive.traindata)
naivebayes
```
#Compare this to the number you obtained in (E).
```{r}
#The value I found above for the naive bayes probability, is .0916 or 9.16% . The number I found in E, was .0905 or 9.05%. The two values are close, only off by .11 %. 
```

